---
title: "Analysis of Electoral Bonds Data"
subtitle: "Published by the ECI on 14-03-2024"
date: "`r Sys.Date()`"
output: html_document
---

```{r load-libraries, include=FALSE}
library(dplyr)
library(readr)
library(kableExtra)
library(tidyr)
options(scipen=999)
```

The datasets were processed from PDF files uploaded on the ECI website. To ensure that the conversion (from PDF to CSV) is error free, we will calculate a few aggregate figures as reported by a few news agencies and compare the final figures. 

```{r read-datasets, include=FALSE}

file_purchaser <- readr::read_csv("../data/purchaser.csv", col_types = cols())
file_encasher <- readr::read_csv("../data/encashment.csv", col_types = cols())

```


## Does the total amount match in both files?

A difference of approx **600 crores** between the two files was also highlighted. 

```{r difference-in-total-amount, include=FALSE}

total_amount_purchaser <- sum(file_purchaser$denomination)
total_amount_encasher <- sum(file_encasher$denomination)

total_diff <- (total_amount_purchaser - total_amount_encasher)/10000000
```

Total amount in the _purchaser_ file is `r round(total_amount_purchaser/10000000)` crores whereas total amount in the _encasher_ file is `r round(total_amount_encasher/10000000)` crores which amount to a difference of INR `r total_diff`(approx) crores.

## Who are the top 10 Purchasers

```{r top-10-purchasers, echo=FALSE}
top_10_purchaser <-
  file_purchaser %>% group_by(purchaserName) %>% summarise(totalAmount = round(sum(denomination)/10000000)) %>% arrange(desc(totalAmount)) %>% head(10)

future_org <- file_purchaser$purchaserName[grepl(pattern = "future",x = file_purchaser$purchaserName,ignore.case = TRUE)] %>% unique()

kable(top_10_purchaser, format = "pipe")
```
_Note: The total amount is in Crores_

In the purchaser file, it looks like the same purchaser has a few different names. 
For e.g. _FUTURE GAMING AND HOTEL SERVICES PR_ has three different names - _FUTURE GAMING AND HOTEL SERVICES PR_, _FUTURE GAMING AND HOTEL SERVICES PRIVATE LIMITED_ & _FUTURE GAMING AND HOTEL SERVICES PVT LTD_. The total amount of bonds purchased by these three entities is equal to `r sum(file_purchaser$denomination[file_purchaser$purchaserName %in% future_org])/10000000` crores.


## Who are the top 10 Encashers

```{r top-10-encashers, echo=FALSE}
top_10_encashers <- file_encasher %>% group_by(politicalParty) %>% summarise(totalAmount = round(sum(denomination)/10000000)) %>% arrange(desc(totalAmount)) %>% head(10)

kable(top_10_encashers, format = "pipe")

```

_Note: The total amount is in Crores_


## Bond denomination distribution by political parties

```{r denomination-distribution-by-parties, echo=FALSE, message=FALSE, warning=FALSE}

denomination_dist <- file_encasher %>% group_by(politicalParty, denomination) %>% summarise(totalBonds=n()) %>% pivot_wider(names_from = denomination,values_from = totalBonds) %>% mutate(totalBonds = rowSums(across(where(is.numeric)),na.rm=TRUE))

denomination_dist_select <-
  denomination_dist[denomination_dist$politicalParty %in% top_10_encashers$politicalParty, ] %>% arrange(desc(totalBonds)) %>% head(5)

kable(denomination_dist_select,format = "pipe")

```

---

Data Source: [Election Commission of India](https://www.eci.gov.in/disclosure-of-electoral-bonds)